Discussion Draft

### 2020-04-03 doc/Discussion.rmd
clean up chromatin org section

Our results from MSM and PWD show that the sc length compaction is independant (of the direction of heterochiasmy. Two strains have have more crossovers in males, but females still have longer sc axis. **This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part of recombination landscape / heterochiasmy**

The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013].

Very few models make predictions for how chromatin organization relates to variation in recombination rates / the recombination landscape.  (the chromatin organization is how chromosomes with the same number of base pairs) are transformed into bottle brush structures with a central axis and most of the genetic material arranged as loops emanating outwards (figure). 
Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume.  This prediction / model might fit broader pattern (such as in At where pollen is the larger cell and has longer axis length) [@cahoonLibuda2019].

Predictions for this trait are  inconclusive (due to lack of cytology measures from both sexes. We argue that the chromatin structure (and axis length) should be considered as an integral parameter when studying recombination landscape.

We also hypothesize that the sex differences in chromatin structure are closely connected to sex differences  in the interference strength and our results mirror those in [@petkov2007] and follow the prediction that shorter-loop will have stronger interference.

According to the tethered loop model of DSB formation (cite schmenti?), the 'real estate' for COs are the chromatin / DNA loops, the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis. **The chromosomes are converted to a 2D space (the axis length of a given chromosome is the limited factor for total CO number)**.  Only one of the models COM aligns with our observed results for sexual dimorphism for axis length. **COM** predicts that axis length will be larger in oocytes due to the larger volume.

>>>>

Our results from MSM and PWD show that the sc length chromatin compaction is decoupled from the direction of heterochiasmy. Two strains have have more crossovers in males, but females still have longer sc axis. **This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part **. These departs from predictions based on empiracal data which nominate chromatin organization (axis length) as the primary driver of heterochiasmy (Petkov).

while sexual dimorphism in chromatin structure is supported by data from mammals literature, especially for mammals [@lynn2002, @gruhn2013]. (but there's generally a lack of data for other systems see cahoon libuda).

There are few models evolutionary models for sex differences in chromatin organization. Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume.  This cell volumne prediction model might fit broader pattern (such as in At where pollen is the larger cell and has longer axis length) [@cahoonLibuda2019].  We argue that the chromatin structure (and axis length) should be considered as an integral parameter when studying recombination landscape.

We also hypothesize that the sex differences in chromatin structure are closely connected to sex differences  in the interference strength and our results mirror those in [@petkov2007] and follow the prediction that shorter-loop will have stronger interference. According to the tethered loop model of DSB formation (cite schmenti?), the 'real estate' for COs are the chromatin / DNA loops, the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis. **The chromosomes are converted to a 2D space (the axis length of a given chromosome is the limited factor for total CO number)**.  Only one of the models COM aligns with our observed results for sexual dimorphism for axis length. **COM** predicts that axis length will be larger in oocytes due to the larger volume.
